\toolnameend: A Tool for Visualizing Patterns of Spreadsheet Function Combinations

### 2016-08-09 doc/paper.tex
Clarify value of spreadsheet novices. Closes #157
Limitation refinement. Closes #158 and closes #148

	We had four participants, labeled P1 through P4. Only P4 had no experience with
	spreadsheets in a professional environment, having used spreadsheets
	intermittently through university and hobbies. P1 through P3 all had varying
	levels of experience with spreadsheets in a professional capacity: P1 for 1
	year, P3 for 3, and P2 for 5. All participants were graduate students at the
	time of study but with some prior experience in industry, and we presented each
	with the same broad question about spreadsheet practice. Though interpretation
	of the phrase ``spreadsheet practice" can certainly vary from participant to
	participant, we readily accepted this interpretive flexibility since we wanted
	a range of views and backgrounds despite the persona.

@@

	\subsection{Study Limitations} \label{subsec:studylimitations} Though we did not start the study by telling
	that the data source came from Enron, two users studied the file names to
	deduce its roots in finance, if not Enron itself. Because of these notorious
	connotations, then, they might have been inclined from this information to
	orient their exploration around finance-related functions, potentially limiting
	their findings or perceived users for the tool. However, one participant also
	said that, without this essential context, they could not have intelligently
	explored a dataset in the first place. Comments like these point out the
	trade-offs of unguided exploration, too: though it might not restrict them to a
	single purpose, it might, at the same time, not afford them any mindset at all
	with which to interpret the data. \par

	Furthermore, the data which users explored was processed and created before we
	had fully ensured that the tool could extract all of the useful data from the
	spreadsheets, meaning some of the counts and patterns in the trees were
	inaccurate. However, we decided that this was not a pressing concern, being
	that we were primarily concerned with whether any interesting reports could be
	made at all, not with whether these reports were 100\% accurate assessments of
	reality. \par

	Throughout the study, we noticed that users misunderstood some of the icons
	and, as such, their early discoveries were inaccurate. For example, P1
	interpreted the numbers on the argument boxes to mean that all of the box's
	children had exactly that many arguments -- that is, clicking on the dark blue
	``1" would return all the single-argument functions instead of the first
	argument of a function -- and P2 initially took the solid lines to mean
	different paths in formula construction rather than different arguments of the
	same function. Though the users naturally did not explicitly report
	misinterpreting the icons, these misunderstandings became clear in their early
	evaluations of the spreadsheet practice. Though we clarified the meanings when
	these misunderstandings became apparent, these events point to a failure in the
	tool to indicate every function clearly.   \par

	\section{Limitations} We began these studies with the intention of evaluating
	our progress towards the goals in Section \ref{goals}. Working from these
	results, we noted many places where the tool aligned well or fell short.

	\subsection{Unbiased Exploration} The first goal -- that of full exploration --
	is threatened by the presence of certain formulae that the tool does not
	display, such as those with nonstandard functions. Because the data collection
	depends on POI's formula parser, it does not allow for any partial parsing of a
	formula; it either processes a formula perfectly or throws it out. As such, the
	visualization does not display anything with syntactical errors or third-party
	functions that POI cannot parse. This poses larger problems, however, when POI
	does not support the parsing of certain functions, such as EOMONTH, preventing
	the healthy growth of those trees. \par

	Other than that, we did avoid designing anything that intentionally evaluated
	formula designs. Nevertheless, \toolname inadvertently emphasizes bad design of
	certain types, such as all of those in Section~\ref{badsmells} marked by their
	visual signatures. Even so, our participants found the discovery of bad smell
	to be the most interesting aspect about it, to the point where they could
	imagine business scenarios in which micromanagers use it to ensure precise
	design parameters across a department's spreadsheets! To design in that
	direction, however, would mean a reevaluation of this design goal, since
	intentionally guiding users toward known smells may guide them away from new,
	unmarked ones.

@@

	A threat to our goals arises from the problem of too much data,
	particularly in the popular functions, like SUM and IF. We realized early that
	these functions would overwhelm the user with options in large datasets, so we
	included features, such as the expansion arrows and the use of generic types
	instead of values, that would manage the variation. Nevertheless, P3 remarked
	that they specifically avoided the most popular functions, anticipating a flood
	of nodes from which they could salvage nothing interesting. P2 corroborated
	this by explaining how they gravitated toward exploring specific examples in
	particular value nodes over navigating the full, crowded tree, but they also
	said that the tool at its best when it compares either every possibility of an
	argument side-by-side or nothing on that branch at all. Despite our
	precautions, balancing the desire to show a lot with the desire not to
	overwhelm remains a challenge. \par

	Some visual cues turned out to be more confusing than helpful, such as
	the index number on each argument node. As discussed in Section~\ref{subsec:studylimitations},
	users often mistook it for other features, like total number of arguments in the
	function. Though we kept this feature present for all studies, future versions
	of the tool would refine features like this to reduce confusion.

@@

	As we found in the case study, the tool excels at making formula examples
	available throughout a tree. However, the benefits described in the previous
	sections imply a hidden complication: if the tool educates through example but
	does not explicitly distinguish between good design and bad design, how can we
	be sure that people will not inadvertently learn bad design from it? In the
	user study, we discussed its use in education and smell detection by bringing
	up the issues surrounding the LOOKUP functions; this discussion, though,
	presupposes a distinguishing eye that knew which examples practiced good design
	and which exhibited the common errors. A beginner might not know the
	difference. Robillard and Deline, in their study, remarked that good API
	examples should demonstrate best practices, but, right now, we have no way to
	assure this with \toolnameend. As with the problems in achieving unbiased
	exploration, correcting for this danger may entail a reevaluation of design
	goals.

>>>>

	We had four participants, labeled P1 through P4. Only P4 had no experience with
	spreadsheets in a professional environment, having used spreadsheets
	intermittently through university and hobbies. P1 through P3 all had varying
	levels of experience with spreadsheets in a professional capacity: P1 for 1
	year, P3 for 3, and P2 for 5. All participants were graduate students at the
	time of study but with some prior experience in industry, and we presented each
	with the same broad question about spreadsheet practice. Though interpretation
	of the phrase ``spreadsheet practice" can certainly vary from participant to
	participant, we readily accepted this interpretive flexibility since we wanted
	a range of views and backgrounds despite the persona. Furthermore, their
	differing levels of expertise let us judge the accessibility of the
	visualization; even if a participant had no formal knowledge of code smells, we
	found it valuable nevertheless if they could intuitively identify bad design.

@@

	\subsection{Study Limitations} \label{subsec:studylimitations}

	In a few ways, these user studies might not fully assess \toolnameposs value.
	For example, though we did not preface the study by revealing the source of the
	files, two users studied the file names to deduce its roots in finance, if not
	Enron itself. Because of these notorious connotations, then, they might have
	been inclined to orient their exploration around financial functions,
	potentially restricting their findings. However, one of these participants also
	said that they could not have intelligently explored a dataset in the first
	place without this essential context. Comments like these point out the
	trade-offs of unguided exploration, too: though it might not restrict them to a
	single purpose, it might, at the same time, not afford them any mindset at all
	with which to interpret the data. \par

	Throughout the study, we noticed that users misunderstood the design
	and, as such, their early discoveries were inaccurate. For example, P1
	thought that the numbers on argument nodes as the number of arguments instead
	of the index of that argument -- that is, clicking on the dark blue
	``1" would return all the single-argument functions instead of the function's
	first argument -- and P2 initially took the solid lines to mean
	different paths in formula construction rather than different arguments of the
	same function. Though these misunderstandings were useful in pinpointing weak
	parts of the tool, their presence affected many of the participant's observations before we
	could clarify, reducing the amount of useful information from these sessions.  \par

	\section{Tool Limitations}

	We began these studies with the intention of evaluating
	our progress towards the goals in Section~\ref{goals}. Working from these
	results, we noted many places where \toolname itself aligned well or fell short.

	\subsection{Unbiased Exploration}

	The goal of \toolname is exploration without bias, gap, or interruption, but 
	this version does not solve everything. First, \toolname cannot reliably
	salvage every formula. When a formula arises that the POI parser cannot handle,
	whether because of typos or third-party functions, the entire formula is lost;
	\toolname gets either the full formula or nothing at all. Second, while we 
	hard-coded nothing that explicitly emphasizes bad design, the nature of the
	visualization draws attention towards certain offenses, like those in
	Section~\ref{badsmells}. While these may be fairly obvious, however, they 
	may distract from subtler, more insidious patterns. \par

	Despite this second point, our participants found the discovery of bad smell
	to be the most interesting aspect about it. To design in that
	direction, however, would mean a reevaluation of this design goal, since
	intentionally guiding users toward known smells may guide them away from new,
	unmarked ones.

@@

	We wanted to show everything, but everything overwhelms. We realized that danger
	early on and included features, such as the expansion arrows and the use of generic types
	instead of values, that would manage the variation. Nevertheless, P2 and P3 remarked
	that they often avoided the most popular functions, anticipating a flood
	of nodes from which they could salvage nothing interesting. P2 corroborated
	this by explaining how they preferred to explore specific examples in
	individuals nodes instead of navigating the full, crowded tree. Despite our
	precautions, balancing the desire to show a lot with the desire not to
	overwhelm remains a challenge. \par

	Some visual cues turned out to be more confusing than helpful, such as
	the index number on each argument node. As discussed in Section~\ref{subsec:studylimitations},
	users often mistook it for other features, like total number of arguments in the
	function. Though we kept this feature present for all studies, future versions
	of \toolname would refine this design.

@@

	As we found in the case study, the tool excels at making formula examples
	available throughout a tree. However, the benefits described in the previous
	sections imply a hidden complication: if the tool educates through example but
	does not explicitly distinguish between good design and bad design, how can we
	be sure that people will not inadvertently learn bad design from it? In the
	user study, we discussed \toolnameposs use in education and smell detection by bringing
	up the issues surrounding the LOOKUP functions. This discussion, though,
	assumes a distinguishing eye that knew which examples practiced good design
	and which exhibited the common errors. A beginner might not know the
	difference. Robillard and Deline, in their study, remarked that good API
	examples should demonstrate best practices, but, right now, we have no way to
	assure this with \toolnameend. As with the problems in achieving unbiased
	exploration, correcting for this danger may entail a reevaluation of design
	goals.
